The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-09-25 paper/ms.tex
paper: add working draft

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

\shortauthors{LGB, KM, JNW}
\shorttitle{Binary Biases}

@@

\title{ Binary Biases \& Toy Transit Surveys }

@@

% \journalinfo{Draft version}
\slugcomment{Memo for internal use}

@@

First, we show that in a universe where all stars have fixed properties
(except that some are twin binaries), the results of a search for planets of a 
given radius and orbital period can be expressed analytically.

@@

Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
show that if we assume every star in this {\it Kepler} survey analog is single, 
we underestimate occurrence rates for a given planet radius and period by 
a factor of $\approx 1.4$, or a percentage error relative to the true 
occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
true}=-28\%$. This is relevant for measuring accurate values of the rate at 
which Earth-like planets exist around Sun-like stars.

@@

Binary companions to single stars introduce a surprising number of biases into 
transit surveys.
For a start, any magnitude limited survey over-represents binaries compared to 
a volume-limited survey -- the familiar Malmquist bias, though in a slightly 
different context.
While transit surveys are ideally signal to noise limited, if the noise 
properties of the survey are dominated by photon counting noise, a signal to 
noise limit becomes equivalent to a magnitude limit.
Thus stars surveyed in transit surveys must also over-represent binaries 
compared to those in a volume-limited sample.

A more straight-forward effect of binarity on transit surveys is that any star 
which is unknowingly a double introduces extra stars about which planets might 
be detectable.
However, this apparent increase in the number of searchable stars is mitigated 
by dilution: the observed transit depths in such systems are reduced by 
contaminating flux from the binary companion. This dilution is especially 
drastic if the planet orbits the secondary, in which case dilution impacts 
detectability the most.
Note that dilution can come from visually associated companions as well as 
gravitationally bound companions.

A consequence of dilution is misinterpretation of transit depths. If the 
planet to star radius ratio is directly taken as the square-root of 
the observed transit depth, derived planetary radii in binary systems will 
systematically too small.
Ciardi et al [2015] explored this issue in depth. They claimed that for the 
case of \kepler, if ones assumes host stars are always single, one will
underestimate radii by an average multiplicative factor of $\approx 1.5$ 
(presumably, compared to the case of even weights for whether the 
primary/secondary gets a given planet).

Finally, binarity may affect transit surveys through actual astrophysical 
processes.
Duquennoy \& Mayor [1991], and alter Raghavan et al [2010] showed that over 
half of solar-like are separated by less than 100\,AU.
One immediate consequence on circumprimary and circumsecondary architectures is 
that dynamical stability limits begin to carve away in realms of planet space 
where planets might otherwise be detectable [e.g., Holman \& Weigert 1999, Li 
et al 2017].
The formation efficiency of various types of planetary systems may also 
be different in binary systems compared to single-star systems.
For instance, the properties of the circumstellar and circumbinary disks 
influence the dynamics of planetesimals [Rafikov, Silsbee];

 the effects of which have been observed [Kraus et al 
2016].
Also, the formation rates of planets, and thus their intrinsic occurrence 
rates, may be influenced by properties of circumprimary and/or circumbinary 
disks.
Although the observations are still ongoing, early results seem to indicate 
that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
not decrease planet occurrence rates [CITE], and may even increase them for 
giant planets [Ngo et al 2017].

@@

Binarity introduces several convoluted effects on transit 
surveys. To disentangle them, imagine the following idealized survey:

\begin{enumerate}
\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
with a detector of area $A$, and known bandpass. Your detector is photon-noise 
limited.
%
\item You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
%
\item You are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. For instance, G2V dwarfs.
%
\item You can only detect your `nominal planet' when you observe signals with
${\rm S/N} > {\rm (S/N)_{min}}$.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux, $F_{\rm min}$.
%
\item However, you opt to observe all the ``points'' on-sky with apparent 
magnitude $m < m_{\rm lim}$, or equivalently with energy flux in your bandpass 
greater than some limit $F > F_{\rm lim}$.
(You will not be able to detect planets around the faintest of 
these, and will need to perform a completeness correction to later derive 
occurrence rates.)
Generally $F_{\rm lim} < F_{\rm min}$, \textit{e.g.}, whenever the original 
magnitude cut defines a ``source catalog'' from which ``searchable stars'' are 
selected, as in the KIC [Batalha et al, 2010, Brown et al 2011] and the TIC 
[Stassun et al 2017].
\end{enumerate}

You get funding, and perform the S/N-limited survey. All planets with 
${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises both single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R$ is a constant across the population 
	of star systems.
	The ratio of the two number densities in a 
	volume-limited sample is the binary fraction\footnote{The binary fraction 
		is equivalent to the multiplicity fraction if there are no triple, 
		quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by these two terms.
	Note that we are taking an limit by assuming the primary and secondary of 
	a binary system host planets at equal rates. If the secondary is 
	non-identical to the primary, it's clearly wrong.	
\end{itemize}

For simplicity, one might wish to imagine the observer correctly 
pre-selects all of the ``searchable stars'', in the style of [Pepper et al, 
2003].
In other words, they would begin by knowing ${\rm (S/N)_{min}}$, and then only 
observe stars for which they would have adequate flux to achieve it -- setting 
$F_{\rm lim} = F_{\rm min}$.
This simplification would naively allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

We do not begin this way because in a stellar population of both singles and 
binaries, the above approach ignores necessary incompleteness for 
binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three different maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which single stars are
  selected. This can be chosen to coincide with the maximum distance out to
  which planets are detectable around single stars.
%
\item $d_{{\rm max},d}^\star$: the maximum distance out to which double stars
  are selected. This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{-1/2} d$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}^\star.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 selected binary stars can yield
  detectable planets.
\end{enumerate}
We develop tools to address this and related completeness effects from the 
outset.

Consider the following questions:
\begin{enumerate}
\item How many single and double star systems, respectively, are in the sample?
Correspondingly, how many stars are in the sample?

\item How many planets are in the sample? (Orbiting single stars, and orbiting 
double stars respectively).

\item What is the true occurrence rate?

\item How many planets are detected?

\item What occurrence rate does astronomer A, who has never heard of binary 
star systems, or completeness, derive for planets of radius $R_p$ and period 
$P$?

\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
effect of binarity (\textit{i.e.} that the sample actually has more stars than 
astronomer A thought) derive? (Astronomer B neglects completeness).

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C did a 
combination of high resolution imaging and RV followup on every candidate, and
correctly classifies the planetary radii in every case.

\item What about astronomer D, who additionally notes the importance of 
completeness?
\end{enumerate}

\subsection{How many stars are in the sample?}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. Then the total number of stars in the sample is
\begin{equation}
N_{\rm stars} = N_s + 2N_d.
\end{equation}

In a magnitude-limited sample in which stars are uniformly distributed in 
volume, the number of stars will be the number density times the volume.
If the volume is taken to be a sphere over which the number density is uniform,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
by definition. The absolute normalization of the number density is a measured 
quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ can also be 
stated in terms of the limiting magnitude $m_{{\rm min}}$,
\begin{equation}
m_{{\rm min}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
min}}}{F_{0}}\right),
\end{equation}
for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
always, everything is implicitly written in a defined bandpass).

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

As a comment on Eq.~\ref{eq:number_systems}, if 
we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either 
$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
the galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.

The number of planets in the sample is
\begin{align}
N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
				  \nonumber \\
			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
\end{align}

The factor of 2 accounts for the fact that there are twice as many stars in 
double star systems.

@@

\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}

\subsection{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
The former is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}} f_{s,{\rm S/N} > {\rm 
(S/N)_{min}}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,{\rm geom}}\equiv f_{s,g}$ is the geometric 
transit probability, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})} \equiv f_{s,c}$ 
is the fraction of these transiting planets that are observed with signal to 
noise greater than the minimum detection threshold (the completeness). 
Analogously,
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample, the geometric transit probability is the same (if we 
have twin binaries -- not if we consider a distribution of secondaries) and the 
completeness term must account for any differences in the signal to noise 
distribution that come from stellar binarity.

\subsubsection{Analytic completeness}
Since the geometric transit probability is known, the only terms we have yet to 
compute are the completeness terms, $f_{i,c}$ for $i\in{\rm \{single,double\} 
}$. We proceed as follows.

>>>>

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{comment}

\shortauthors{LGB, KM, JNW}
\shorttitle{Binary Biases \& Toy Transit Surveys}

@@

\title{ The Effects of Stellar Binarity on Transit Survey Occurrence 
Rates }

@@

% \journalinfo{Draft version}
\slugcomment{Draft version}

@@

\begin{comment}
First, we show that in a universe where all stars have fixed properties
(except that some are twin binaries), the results of a search for planets of a 
given radius and orbital period can be expressed analytically.

@@

Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
show that if we assume every star in this {\it Kepler} survey analog is single, 
we underestimate occurrence rates for a given planet radius and period by 
a factor of $\approx 1.4$, or a percentage error relative to the true 
occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
true}=-28\%$. This is relevant for measuring accurate values of the rate at 
which Earth-like planets exist around Sun-like stars.
\end{comment}

@@

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
of stellar $(M_\star, R_\star, L_\star)$ phase-space.
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that appear to be searchable for the 
desired type of planet, and also of the desired stellar type.  These 
``searchable stars'' are the points on the sky for which they think planets 
are observed with 100\% detection efficiency. Accounting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

\begin{equation}
\Gamma_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
\end{equation}

There are many potential pitfalls.  Some genuine transit signals can be missed
by the detection pipeline.  Some apparent transit signals are spurious, from
noise fluctuations, failures of `detrending', or instrumental effects.  Stars
and planets can be misclassified due to statistical and systematic errors in
the measurements of their properties.  Poor angular resolution causes false
positives due to blends with background eclipsing binaries. {\it Et cetera}.

Here we focus on problems that arise from the fact that stars of the desired
type often exist in binaries. 
We assume for simplicity that all binaries in the transit survey are 
spatially unresolved.

One immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets of the desired 
type may differ between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval $\Delta M_\star, 
\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
over stellar multiplicity.
Otherwise, one must discuss ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

Outside of bonafide astrophysical differences, there are observational biases.
A given apparently-searchable star may truly be a single star of the desired 
type. If not, and ignoring higher order multiples, one of the following must 
be true:
\begin{enumerate}	
\item The apparently-searchable star is a binary system, with two searchable 
stars of the desired type. $N_\star$ is under-counted by one for every such 
system.
%
\item The apparently-searchable star is a binary system, with one searchable 
star of the desired type. 
The searchable star of the desired type could be the primary or the secondary.
$N_\star$ is correctly counted for every such system, though there may be 
systematic errors in estimates of stellar parameters.
%
\item  The apparently-searchable star is a binary system, in which neither of 
the components is a searchable star, but at least one is of the desired type. 
$N_\star$ is over-counted by one for every such system.
\end{enumerate}

One might also imagine an apparently-searchable star which in no stellar 
component is of the desired type.
For single stars, this may simply be from erroneous stellar parameters.
In binaries, one might imagine the combined light of two stars somehow making 
them resemble a single star of the desired type.  We will 
not consider errors of this type.  We will assume that all the 
apparently-searchable stars are either single stars of
the desired type, or binaries in which either the primary or else both 
components are of the desired type.

The above enumerates possible errors when counting $N_\star$.
When tallying $N_{\rm det}$, the detected signals from planets of the desired 
type, neglecting binarity introduces the following error cases:
\begin{enumerate}
\item 
A detected signal is from a star of undesired type (by the above assumptions, 
the secondary of a binary system), from a planet of undesired type, and is 
incorrectly counted as a planet of desired type.
%
\item
A detected signal is from a star of desired type (the primary or secondary of 
a binary systems), from a planet of undesired type, and is incorrectly counted 
as a planet of desired type.
This and the case above occur when the host star parameters are miscalculated 
({\it e.g.,} the host star is the faint secondary, but is assumed to be 
primary), or when the constant diluting light from the binary companion is not 
corrected.
$N_{\rm det}$ should be lowered by 1 for every such case.
%
\item
A detected signal is from a star of desired type, from a planet of desired 
type, but incorrectly counted as planet of undesired type.
$N_{\rm det}$ should be raised by 1 for every such case.
\item
An {\it undetected} signal from a planet of the desired type, around a star of 
the desired type, was not detected because of dilution from the companion 
(because the SNR floor of the survey is set for single stars). 
This happens during case \#3 in $N_\star$-counting errors (the 
apparently-searchable star is a binary system, in which neither component is 
in fact searchable). If $N_\star$ is corrected for this, {\it i.e.,} the 
system is correctly counted as ``not searchable'', no correction to $N_{\rm 
det}$ is needed.
\end{enumerate}

Errors (1)-(3) in calculating $N_{\rm det}$ above broadly fall under ``radius 
misclassification'', while error (4) falls under ``completeness 
miscalculation''.
Note that in the above list, a planet cannot be both ``of desired type'' and 
``orbiting a star of undesired type''. We ignore this type of error for 
analytic convenience, but \texttt{reintroduce it later during numerics}.

Our approach is step-by-step starting from a very simple scenario.
\begin{itemize}
\item We consider a universe consisting of only the desired types of stars and
planets.  We give analytic expressions for eta(apparent)/eta(true) taking
into account all applicable errors.
%
\item We allow for a distribution of the mass ratio $dn/dq \propto q^\gamma$, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
We give analytic expressions for eta(apparent)/eta(true) taking into 
account errors of type 2.
%NOTE: empirical L(M) maybe OK for numerics -- not for analytics.
%
\item We allow for a nonzero tolerance $\Delta R_p$ in the planet radius, and 
tolerance $\Delta L_\star$ in stellar luminosity, to qualify as "desirable".  
We recalculate eta(apparent)/eta(true).
(LB: I think this has to become numerical to actually evaluate the occurrence
rate correction factor, $X_\Gamma$.)
%
\item We assume the planets have a power-law radius distribution, $d\eta/dR_p
\propto r^\alpha$.  We calculate the apparent $d\eta/dR_p$ taking into account 
errors of type 1-3.  We do so for stars of the desired type within some 
tolerance dL.
%
\item Everything up to now is more-or-less analytic, perhaps supported by 
numerical checking or numerical integration. At the end we summarize the 
radius-dependence of the occurrence rate shifting.
"Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between"

\end{itemize}

We then dig in, w/ numerics, to what happens with different assumed
completeness vs Rp functions.
We note that the the claimed HJ occ rate discrepency btwn RV \&
transit surveys gets much weaker when accounting for binarity.
This is b/c the more believable occ rate papers do not heavily rely on their
completeness calibration -- they know they're complete.
This is a much more understandable regime than the wonky completeness estimate 
dependent zone of eta-Earth for Kepler.

Finally, for numerics, we simulate a sample of 20,000 
apparently-searchable-stars (or however many "stars" appear to be searchable
for Earthlike planets in the Kepler field).  Basically this means a magnitude
limited survey of 20,000 ``stars'' that are either single or are binaries with
the primary of the desired type.  The sample has a realistic distribution of
binaries (occurrence and luminosity ratio).  We calculate 
eta(apparent)/eta(true).  We can also play with errors of type 4 in this
numerical model (i.e. we let the relative occurrence rates between primaries 
and secondaries vary).  (It may not worth doing so for the analytic 
work-- though it's pretty simple. The way to play with this is just present 
both limits, and then also present a few weight factors (Furlan+17 do a good 
job).

We discuss ``how do we interpret Fulton+ 2017?''
That paper did (or claims to have done) a good job at weeding out numerator
errors. If true, means $\Gamma(R_{\rm p})$ has the right shape, and just the 
wrong normalization. Important if we ever manage to convince ourselves 
``brightest transit hosts'' could work.

In passing, at the end, we mention how this changes attempts at eta-Earth 
measurements.

\begin{comment}
UNSOLVED PROBLEMS W/ THE NUMERICAL MODEL-
* currently have not implemented self-consistent (L,M,R) relation between
primaries and secondaries.
Wang et al (2015) avoided that problem by doing a similar empirical thing to
what I currently have implemented. But I do not think that was a good paper.
Ciardi et al (2015) have an approach that was Kepler-specific -- they used
the same Dartmouth isochrones used for KIC stellar parameters.
But they did not conserve $dn/dm_x$, for $m_x$ apparent mag in band x.
I think the latter choice might be better for Kepler-specific discussion.
But it's less-nice if we want this to be general enough for e.g., occ HJ
rates w/ TESS.

My main gripe w/ taking the 150k KIC stars \& adding binaries is that less
is known about those parameters and biases. Using e.g., Galaxia has the major
benefit of you understand what's been cooked in much better (it's minimal).

I think if we want to talk about Kepler, it's the best possible approach
though.

If we want to talk about Fulton+17 we can go further, and apply the CKS cuts.

Some questions based on the current draft:

- Why should we ever consider $F_{lim} < F_{min}$?

LB: in practice, surveys look at many stars for which they could not detect
Rp=Re planets, but could detect Rp=10Re planets.  One possible error is
miscounting which stars are which (i.e. getting completeness corrections
wrong). So this was more for plausbility than b/c it's immediately needed in
the analytics. I can cut it.

- Why do we need to consider the distribution of S/N, as opposed to simply the
limiting distance for detection?

LB: for "model 1" (fixed stars, fixed planets, fixed light ratio binaries)
you're correct, it's not immediately necessary.

For this model, I did it as a way to think through the completeness effects
systematically -- i.e. how for the twin-binary case, you get a completeness
fraction of 1/8, though your limiting distance is sqrt(2) times bigger. (Even
when your completeness for the singles is 100%).

If we go the path of completing "model 2"'s analytics (vary light ratio, fixed
primary luminosity, fixed planet), I think this formulation would become
necessary in order to actually evaluate the completeness fractions.

(I never completed these analytics b/c of the L(M) mess + confusion about
assumed stellar masses+radii. Power law L(M) helps this... I think there needs
to be an extra assumption about the stellar mass+radius change though that we
have not spelled out)

- I don't think the bandpass, effective area, photon count rate, etc.  are
needed, either. All that matters is the limiting distance.

Are needed for what? 
In Model 1, for analytic expressions of the occurrence rate, I agree.
For analytic expressions of the number of detected planets, I disagree.

For Model 2, I'm not sure because I got bogged down. I suspect that for
occurrence rate errors, you are correct.

The analytic expression I gave for the occurrence rate depends on this "xi"
factor, which needs to incorporate an assumption about how Astronomer A
misestimates the completeness (relatedly, how he misestimates Rstar).  I never
specified that assumption.

- In Eq. 4, is this a traditional definition of BF?  I would have defined it as
nd/(ns+nd) or perhaps 2*nd/(ns+ 2*nd).  Let's make sure we are using
traditional terminology, and also interpreting the literature correctly, i.e.,
is the right number really 0.45?

Oops. nd/(ns+nd) = BF = MF (if assuming all multiple systems are binaries).
This minor change propagates.

Touching up on correct terminology, from Duchene \& Kraus 2013, Sec 2.1:

"Frequency of multiple systems" (MF)
vs
"Companion frequency" (CF; average \# of companions per target, can exceed 
100\%)

my understanding of the terms:

survey 1: {*, **, *, ***, *, *}

6 systems.
2 are multiple.
MF = 2/6 = 1/3
CF = (0+1+0+2+0+0)/6 = 3/6 = 1/2

survey 2: {**, **, ***, ****, *}

5 systems
4 are multiple
MF = 4/5
CF = (1+1+2+3+0)/5 = 6/5

DK2013 Sec 3.1.1 give (citing Raghavan+ 2010)
MF^{MS}_{1-1.3Msun} = 50 \pm 4%
MF^{MS}_{0.7-1Msun} = 41 \pm 3%

They also give
CF^{MS}_{1-1.3Msun} = 75 \pm 5%
CF^{MS}_{0.7-1Msun} = 56 \pm 4%

I understand the first pair of numbers to mean:

In a volume-limited survey (of the stellar neighborhood), if you select systems
for which the primary is from 1-1.3Msun, 50% of them will be multiple.

I take the second pair of numbers to mean:

In a volume-limited survey (of the stellar neighborhood), if you select systems
for which the primary is from 1-1.3Msun, each system will on average have 0.75
companions.

So YES, need to correct this throughout.
\end{comment}

@@

Stellar binarity introduces convoluted effects on transit surveys.
To disentangle them, imagine the following idealized survey.

You observe the entire sky for some duration with a detector of known 
area and bandpass. Your detector is photon-noise limited.
You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
You are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. For instance, G2V dwarfs.
You can only detect your desired type of planet when you observe signals 
with ${\rm S/N} > {\rm (S/N)_{min}}$.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux required for detection, $F_{\rm min}$.

You get funding, and perform the S/N-limited survey. 
You observe all the ``points'' on-sky with apparent 
magnitude $m < m_{\rm min}$, or equivalently with energy flux in your bandpass 
greater than some limit $F > F_{\rm min}$.
All planets with ${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are 
not.

You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twin systems.
	The ratio of the number density of binary systems and the number density of 
	all systems in a volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\begin{itemize}
	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have planets of radius $R_p$, with orbital period $P$.
	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
	systems have planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets at 
	equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
	\end{itemize}
\end{itemize}

The assumption of [Pepper et al, 2003] in a very similar context was that the 
observer correctly pre-selects all of the ``searchable stars''.
This simplification would allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

The above approach ignores incompleteness for binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three distinct maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
	of planet is detectable around single stars.
	These are the selected ``searchable, single stars''.
%
\item $d_{{\rm max},d}$: the maximum distance out to which double stars
  are selected (down to the same minimum flux as that defining $d_{{\rm 
  max},s}$). 
  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{-1/2} d$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
  are in fact binary systems can yield detectable planets.

\end{enumerate}

We note in passing that we are considering a SNR-limited survey, which in this 
case is the same as a magnitude-limited one.
The observer might prefer to perform a completeness-limited survey, which in 
this case would be the same as a volume-limited one with limiting distance
$d_{{\rm max},d}^{\rm p}$.
This change would likely improve interpretability.

With the stage set, we ask:

@@

The ``true occurrence rate'' $\Gamma_t$ is the average number of planets of 
the desired type per star of the desired type:
\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + (1+w_{d2}) \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}
where $N_s$ is the number of single star systems, and $N_d$ the number of 
double star systems. 
The factor of 2 in the total number of stars accounts for the fact that there 
are twice as many stars in double star systems.

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
Note that these latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~we count the $\approx 25\%$ of multiple 
systems that are higher order multiples as doubles.
